Viscous Erosion in Porous Media

### 2018-02-27 docs/methods/erosion_methods.tex
NM
Adding GMRES count table

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Convergence test: We run single-body erosion simulations to a stopping time of $t_s =$ 1E-2 (56\% of the vanishing time) and measure the $L^2$-difference in shapes obtained with successive temporal resolutions. The tests confirm the method to be second-order accurate in time.
}
\vspace{0.3 pc}
\label{convtab}
\begin{tabular}{l l l l}
\hline
\hspace{0.0pc} $\Delta t/t_s$
\hspace{0.5pc} & error 
\hspace{0.5pc} & order
\hspace{0.5pc} & CPU time \\
\hline
%
1/50		& 1.68E-5		& --		& 53 secs     	\\
1/100	& 4.84E-6		& 1.79	& 1.8 mins   	\\
1/200	& 1.30E-6		& 1.90	& 3.5 mins  	\\
1/400	& 3.36E-7		& 1.95	& 6.9 mins  	\\
1/800	& 8.54E-8		& 1.98	& 14 mins   	\\
1/1600	& 2.15E-8		& 1.99	& 28 mins  	\\
1/3200	& 5.40E-9		& 1.99	& 56 mins    	\\
1/6400	& --			& --		& 111 mins	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Parameters in this test are: 01circ256, nits = 7, dt0 = 1e-4, tfin = 1e-2
% epsfac = 10, sigfac = 10, fixarea = 0, fixprdop = 0

\subsection{Multiple-body simulations}
With the numerical method validated, we are now ready to simulate the erosion of multiple bodies.

Possible list:
\begin{itemize}
\item Show multiple body simulation with fixed flow rate
\item Then simulation with fixed pressure drop
\end{itemize}

\vsp{30}

% NEWPAGE
\newpage

Runs to do
\begin{itemize}
  \item Single body
  \begin{itemize}
    \item Drag and resistivity
  \end{itemize}
  \item $\bigO(5)$ bodies
  \begin{itemize}
    \item Drag and resistivity
    \item Anisotropy
    \item Long channels between flat faces
    \item Vorticity
  \end{itemize}
  \item $\bigO(50)$ bodies
  \begin{itemize}
    \item Drag and resistivity
    \item Anisotropy
    \item Vorticity
  \end{itemize}
\end{itemize}

>>>>

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Convergence test: We run single-body erosion simulations to a stopping time of $t_s =$ 1E-2 (56\% of the vanishing time) and measure the $L^2$-difference in shapes obtained with successive temporal resolutions. The tests confirm the method to be second-order accurate in time. The last column shows the measured CPU time which increases linearly as the time-step is refined.
}
\vspace{0.3 pc}
\label{convtab}
\begin{tabular}{l l l l}
\hline
\hspace{0.0pc} $\Delta t/t_s$
\hspace{0.5pc} & error 
\hspace{0.5pc} & order
\hspace{0.5pc} & CPU time \\
\hline
%
1/50		& 1.68E-5		& --		& 53 secs     	\\
1/100	& 4.84E-6		& 1.79	& 1.8 mins   	\\
1/200	& 1.30E-6		& 1.90	& 3.5 mins  	\\
1/400	& 3.36E-7		& 1.95	& 6.9 mins  	\\
1/800	& 8.54E-8		& 1.98	& 14 mins   	\\
1/1600	& 2.15E-8		& 1.99	& 28 mins  	\\
1/3200	& 5.40E-9		& 1.99	& 56 mins    	\\
1/6400	& --			& --		& 111 mins	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Parameters in this test are: 01circ256, nits = 7, dt0 = 1e-4, tfin = 1e-2
% epsfac = 10, sigfac = 10, fixarea = 0, fixprdop = 0

\subsection{Multiple-body simulations}

With the numerical method validated, we are now ready to simulate the erosion of multiple bodies. We make one important conceptual change to our model. For a realistic porous medium composed of many solid bodies, the condition of a fixed incoming flow rate $\umax$ is somewhat artificial. A more realistic condition is that the pressure difference from one end to another is held fixed. For example, in groundwater flow, this condition would correspond to a recharge region of given hydraulic head connecting to a discharge region of another hydraulic head.

We therefore impose the condition of a fixed pressure difference by controlling the incoming flow-rate $\umax$ at each time-step. This is achieved by first computing the Stokes flow with $\umax = 1$ prescribed, then measuring the pressure at a location upstream ($x = -2$) and downstream ($x=2$) of the eroding bodies. These upstream and downstream pressures are given by
\begin{align}
& p_{-} = \frac{1}{2} \int_{-1}^{1} p(-2,y) \, dy \, , \\
& p_{+} = \frac{1}{2} \int_{-1}^{1} p(+2,y) \, dy \, .
\end{align}
We then rescale the computed density function by the pressure difference
\begin{align}
\eeta \leftarrow \frac{\eeta}{(p_{-} - p_{+})} \, ,
\end{align}
This rescaling maintains a {\em fixed} pressure difference. Because the Stokes equations are linear, such a rescaling is permissible and does not alter the accuracy of our method. Furthermore, rescaling by a different factor at each time-step is also permissible because the Stokes equations are memoryless (i.e.~there is no time derivative present in the PDEs).

Briefly, to confirm that the rescaling does not alter accuracy, we show HERE a convergence test for 20 bodies with the condition of a fixed pressure drop.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.9 \textwidth]{./figs/06bod.pdf}
\caption{\label{fig:50bodies} 6 bodies
}
%A time lapse of 50 bodies eroding in a viscous flow.  The boundary conditions are a constant pressure drop that creates a flow from the left to right.  The rate of erosion is proportional to the shear stress which is equivalent to the vorticity evaluated on the bodies.  The colors represent the vorticity, so regions with large vorticity (in magnitude) are regions where the flow will erode the fastest.
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1

\vsp{10}

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{GMRES Iterations
}
\vspace{0.3 pc}
\label{itertab}
\begin{tabular}{c c | c c}
\hline
\hspace{0.0pc} Number of bodies
\hspace{0.5pc} & GMRES iterations 
\hspace{0.5pc} & aaa
\hspace{0.5pc} & bbb \\
\hline
%
10	& 223	& a	& b	\\
15	& 292	& a	& b	\\
20	& 363	& a	& b	\\
30	& 496	& a	& b	\\
40	& 676	& a	& b	\\
50	& 991	& a	& b	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% In these tests I always use 128 points per body.

\vsp{10}
Possible list:
\begin{itemize}
\item Show multiple body simulation with fixed flow rate
\item Then simulation with fixed pressure drop
\item GMRES counts
%
\item $\bigO(5-10)$ bodies
	\begin{itemize}
	\item Area and Drag
	\item Long channels between flat faces
	\end{itemize}
\item $\bigO(50)$ bodies
	\begin{itemize}
	\item Area and Drag
	\end{itemize}
\end{itemize}
